---
title: "Do Annual Salaries Of Registered Nurses Today Differ Than 20 Years Ago?"
execute:
  echo: false
---

```{r, include = FALSE}
library(tidyverse)
library(GGally)
```

```{r, include = FALSE}
nurses <- read_csv("https://raw.githubusercontent.com/36-SURE/36-SURE.github.io/main/data/nurses.csv")
```

### Glimpse of Data:
```{r}
### Live look at the data and columns

nurses %>% glimpse()
```


```{r, include = FALSE}
### Better view of column names
nurses %>% names()
```


```{r, include = FALSE}
### Select Necessary Features in data + remove null value rows
cleaned_nurses <- nurses %>% select(State, Year, `Total Employed RN`, `Annual Salary Avg`, `Annual 10th Percentile`: `Annual 90th Percentile`) %>%
  na.omit()
```

### Correlation + Decision Process Of Feature Selection:

```{r, warning=FALSE}
nurses |> 
  select(-State, - `Hourly Wage Avg`, -`Hourly Wage Median`, -`Hourly 10th Percentile`: -`Yearly Total Employed (State)_Aggregate`) %>%
  ggpairs(title = "Correlations Amongst Several Features", 
          upper = list(continuous = wrap("cor"))) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    strip.text = element_text(size = 10),
    axis.text.x = element_text(size = 1, angle = 45, hjust = 1),
    axis.text.y = element_text(size = 5)
  )
```

Notice, that the Annual Salary Avg and Year is the 2nd highest correlation but #1 most relevant -\> since Annual Salary Median and Annual Salary Mean are very highly correlated due to the terms median and mean being very similar to each other: Mean is middle 50% and mean is the average so we expect median to be very close to mean correlation

### Let's Take a Closer Look at the Average Annual Salary per Year Amongst the US States

```{r}
cleaned_nurses %>%
  group_by(Year) %>%
  summarize(average_salary = mean(`Annual Salary Avg`)) %>%
  ggplot(mapping = aes(x = Year, y = average_salary)) +
  geom_point() +
  geom_line() +
  labs(title = "Average US Salary Among The Years")
```

### Now that we know why I chose `Year` and `Annual Salary Average`, let's create a plot with the distribution of `annual salary average` among the US and facet it with `year`.

```{r}
cleaned_nurses %>%
  group_by(Year) %>%
  ggplot(aes(State, `Annual Salary Avg`)) +
  geom_point(alpha = 0.8, color = "chocolate4") +
  facet_wrap(~ Year, nrow = 3) +
  labs(
    title = "Distribution of Average Annual Salary Across the Years",
    x = "State",  # Keep x-axis label
    y = "Average Annual Salary"
  ) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10),
    axis.text.y = element_text(size = 5.5)
  )

```

Note: Scales change as we increase the Years -\> Y Max goes up.

```{r, include = FALSE}
new_cleaned_nurses <- cleaned_nurses %>%
  group_by(State, Year) %>%
  summarize(average_salary = mean(`Annual Salary Avg`), 
            tenth_percentile = mean(`Annual 10th Percentile`),
            twenty_fifth_percentile = mean(`Annual 25th Percentile`),
            seventy_fifth_percentile = mean(`Annual 75th Percentile`),
            ninety_percentile = mean(`Annual 90th Percentile`)
            )


# Assuming salaries_nurses is your data frame and it has columns for the different percentiles
cleaned_nurses_long <- new_cleaned_nurses %>%
  pivot_longer(
    cols = c(tenth_percentile, twenty_fifth_percentile, seventy_fifth_percentile, ninety_percentile),
    names_to = "percentile",
    values_to = "salary"
  )
```

### Same Graph, but with percentiles added on:

```{r}
cleaned_nurses_long %>%
  ggplot(aes(State, salary, color = percentile)) +
  geom_point(alpha = 0.6) +
  facet_wrap(~ Year, nrow = 5) +
  labs(
    title = "Average Annual Salary Comparasion Of The 5 Recent Years VS First 5 Years",
    x = "State",  # Keep x-axis label
    y = "Average Annual Salary (Same scale on purpose)"
  ) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    #plot.title = element_text(hjust = 0.3),
    strip.text = element_text(size = 10),
    axis.text.y = element_text(size = 5.5)
  )

```

```{r, include = FALSE}

# Assuming salaries_nurses is your data frame and it has columns for the different percentiles
short_cleaned_nurses_long <- new_cleaned_nurses %>%
  filter(Year <= 2002 | Year >= 2016) %>%
  pivot_longer(
    cols = c(tenth_percentile, twenty_fifth_percentile, seventy_fifth_percentile, ninety_percentile),
    names_to = "percentile",
    values_to = "salary"
  )
```

### Let's Compare The Recent 5 Years with The First 5 Years of The Data Set and Split The Salaries Down Into Percentiles To Have A Better Look:

```{r}
short_cleaned_nurses_long %>%
  #group_by(State) %>%
  ggplot(aes(State, salary, color = percentile)) +
  geom_point(alpha = 0.8) +
  facet_wrap(~ Year, nrow = 2) +
  labs(
    title = "Average Annual Salary Comparasion Of The 5 Recent Years VS First 5 Years",
    x = "State",  # Keep x-axis label
    y = "Average Annual Salary (Same scale on purpose)"
  ) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text = element_text(size = 10),
    axis.text.y = element_text(size = 5.5)
  ) 

```

### Hey, there are too many scatterplots and facets, please just put it all into 1 plot...

```{r}
cleaned_nurses_long %>%
  ggplot(mapping = aes(x = Year, y = salary, color = percentile)) +
  geom_point() +
  labs(
    y = "Average Salaries",
    title = "Average Salaries Over The Years Broken Down Into Percentiles"
  ) 

```


```{r, include = FALSE}
### Let's now add the different salary percentiles and see how they line up and are placed amongst the data

salaries_nurses <- cleaned_nurses %>%
  group_by(Year) %>%
  summarize(average_salary = mean(`Annual Salary Avg`), 
            tenth_percentile = mean(`Annual 10th Percentile`),
            twenty_fifth_percentile = mean(`Annual 25th Percentile`),
            seventy_fifth_percentile = mean(`Annual 75th Percentile`),
            ninety_percentile = mean(`Annual 90th Percentile`)
            )
```

### An Even Better Plot To Visualize The Same Thing

```{r}

ggplot(cleaned_nurses_long, aes(y = average_salary, x = salary, color = percentile)) +
  geom_point(size = 3) +
  scale_color_manual(
    values = c(
      "tenth_percentile" = "red",
      "twenty_fifth_percentile" = "royalblue",
      "seventy_fifth_percentile" = "gold",
      "ninety_percentile" = "yellow"
    ),
    labels = c(
      "twenty_fifth_percentile" = "25th Percentile",
      "tenth_percentile" = "10th Percentile",
      "seventy_fifth_percentile" = "75th Percentile",
      "ninety_percentile" = "90th Percentile"
    )
  ) +
 
  labs(
    title = "An Even Clearer Distribution of Average Annual Salary Across the Years",
    x = "Percentiles Of Salaries Over The Years",
    y = "Average Annual Salary",
    color = "Percentile"
  ) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.title = element_text(hjust = 0.5),
    strip.text = element_text(size = 10),
    axis.text.y = element_text(size = 5.5)
  ) +
  coord_equal() +
  ggthemes::scale_color_tableau()
  


```
